**Density Dependence**

### 2023-02-15 Rmd/report whb.Rmd
report.R&Rmd for WHB

source("../R/dd.R")
source("../R/get_dropbox.r")
source("../R/pm.R")

dropboxdir<-try(file.path(get_dropbox(), "pelagics"))

@@

# mystk     <- "whb"; mystkname <- "Blue whiting"
# mystk     <- "her"; mystkname <- "North Sea herring"
mystk     <- "mac"; mystkname <- "Northeast Atlantic mackerel"; maxyear <- 2050

@@

# Evaluating impacts of Density Dependent processes on `r mystkname` MSE

**L. Kell, M.A. Pastoors

@@

## Introduction

## OM conditioning

Stock assessments with SAM, Density dependent SAM and Density dependent VPA

```{r, mac.data, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}

load(file.path(dropboxdir,"data/om/mac.RData"))
combY <- as.data.frame(pmYear(window(mac, start=2001)),drop=T) %>% mutate(stk=mystk, type="SAM")

par=FLPar(c(m1=0.15/(0.2^-0.288),m2=-0.28,bmsy=3500000,b=0.3,k=30.1, w50=0.2))

stock.wt(   mac)[1]=0.01
catch.wt(   mac)[1]=0.01
landings.wt(mac)[1]=0.01
discards.wt(mac)[1]=0.01

om_orig=ddFn2(mac,par); 
combY <- bind_rows(combY,
                   as.data.frame(pmYear(window(om_orig, start=2001)),drop=T) %>% mutate(stk=mystk, type="DD SAM"))

om=om_orig+VPA(om_orig)
om=ddFn2(om,par)
combY <- bind_rows(combY, 
                   as.data.frame(pmYear(window(om, start=2001)),drop=T) %>% mutate(stk=mystk, type="DD VPA") )

combY %>% 
  filter(qname %in% c("catch","ssb","rec","fbar")) %>% 
  ggplot(aes(x=year, y=median)) +
  theme_bw() +
  geom_ribbon(aes(fill=type, ymin=lowq, ymax=uppq, xmin=year, xmax=year), alpha=0.3) +
  geom_line(aes(colour=type)) +
  facet_grid(qname~., scales="free_y")

@@

@@

Forward projections with F=fmsy (fwd), 0.8xfmsy?, 1.0xfmsy, 1.2xfmsy?

```{r, mac.data2, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}

sr=fmle(as.FLSR(om,model="bevholt"),control=list(silent=TRUE))
eq=FLBRP(om,sr=sr)

om=fwdWindow(om,end=maxyear,eq)
f =fbar(om)[,ac(2021:maxyear)]%=%refpts(eq)["msy","harvest"]
om=fwd(om,fbar=f,sr=eq)

om2=om
om3=om
om4=om

for (iYear in ac(2021:maxyear)){
  om2=ddFn(iYear,om2,par)
  om2=fwd(om2,fbar=f[,iYear]*0.8,sr=eq)

  om3=ddFn(iYear,om3,par)
  om3=fwd(om3,fbar=f[,iYear],sr=eq)

  om4=ddFn(iYear,om4,par)
  om4=fwd(om4,fbar=f[,iYear]*1.2,sr=eq)
  }

plot(FLStocks("fwd"=om,"0.8"=om2,"1.0"=om3,"1.2"=om4))

@@

## Simulation results

## Discussion and conclusions

ICES
Fmsy
HCR1 (no bounds)

```{r, runWhb1, echo=FALSE, fig.asp=1.15, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}

myvar     <- c("catch","cCum", "ssb","fbar", "rec")
scenarios <- c("ICES", "Fmsy", "HCR1")
rp        <- refs  %>% 
  filter(stk == mystk, refpoint %in% c("fmsy", "blim", "msybtrigger")) %>% 
  mutate(qname = factor(qname, levels=myvar))
ices      <- combY %>% 
  filter(stk == mystk, qname %in% c(myvar),  .id %in% "ICES",  year %in% 2000:2021) %>% 
  dplyr::select(- .id) %>% 
  mutate(qname = factor(qname, levels=myvar))
df        <- combY %>% 
  filter(stk == mystk, qname %in% c(myvar),  .id %in% scenarios, year %in% 2000:2021) %>% 
  mutate(.id = factor(.id, levels=scenarios)) %>% 
  mutate(qname = factor(qname, levels=myvar))
iters        <- combYI %>% 
  filter(stk == mystk, qname %in% c(myvar),  .id %in% scenarios, year %in% 2000:2021, iter <= 5) %>% 
  mutate(.id = factor(.id, levels=scenarios)) %>% 
  group_by(stk, .id, iter) %>% 
  mutate(perc_diff = (data-lag(data))/lag(data)) 
# box <- 
#   bind_rows(
#     df   %>% dplyr::select(.id, qname, stk, data=median),
#     ices %>% dplyr::select(qname, stk, data=median) %>% mutate(.id="ICES")
#   ) %>% 
#   group_by(qname, stk) %>% 
#   summarise(max = max(data, na.rm=TRUE), min = min(data, na.rm=TRUE)) %>% 
#   mutate(min = ifelse(min<0, min, 0)) %>%

df %>% 
  ggplot(aes(x=year, y=median)) +
  theme_bw() +
  theme(panel.grid.minor = element_blank(),  
        panel.border     = element_rect(linetype = 1, colour="black", fill=NA))+

  # vline at 2010
  geom_vline(xintercept = 2010, colour="darkgray", linewidth=1) +

  # iters
  geom_line(data=iters,
            aes(y=data, group=iter),
            linewidth=0.4, colour="darkgray") +

  # median and ribbon
  geom_line(aes(y=median),
            linewidth=1, colour="red") +
  geom_ribbon(aes(xmin=year, xmax=year, ymin=lowq, ymax=uppq),
              alpha=0.2, fill="red") +

  # ICES
  geom_line(data= ices,
            aes(year,median),
            col="blue", linewidth=1, linetype=2, inherit.aes=FALSE) +

  # refpoints
  geom_line(data=rp,
            aes(x=year, y=data, group=reftype, col=reftype, linetype=reftype),
            linewidth=1) +
  scale_linetype_manual(values = c("longdash","dashed"), breaks = c("lim", "msy") ) +
  scale_colour_manual  (values = c("black","darkgray"), breaks = c("lim", "msy") ) +

  expand_limits(y=0) +
  facet_grid(qname ~ .id, scales="free_y")

@@

**Figure `r mystk`.2**  `r mystkname`. Comparison of different HCR’s without TAC bounds

@@

#### `r mystkname`: comparing different HCR’s with and without TAC bounds

Fmsy
HCR0 (no bounds)
HCR1 bnd (-25%/+20% TAC change bounds apply when stock is above MSY Btrigger)
HCR2 bnd (-25%/+20% TAC change bounds apply when stock is above Blim)

```{r, runWhb2, echo=FALSE, fig.asp=1.15, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}

myvar     <- c("catch","cCum", "ssb","fbar")
scenarios <- c("Fmsy", "HCR1",  "HCR1 bnd", "HCR2 bnd")
rp        <- refs %>% 
  filter(stk == mystk, refpoint %in% c("fmsy", "blim", "msybtrigger")) %>% 
    mutate(qname = factor(qname, levels=myvar))
ices      <- combY %>% 
  filter(stk == mystk, qname %in% c(myvar),  .id %in% "ICES",  year %in% 2000:2021) %>% 
  dplyr::select(- .id) %>% 
  mutate(qname = factor(qname, levels=myvar))
df        <- combY %>% 
  filter(stk == mystk, qname %in% c(myvar),  .id %in% scenarios, year %in% 2000:2021) %>% 
  mutate(.id = factor(.id, levels=scenarios)) %>% 
  mutate(qname = factor(qname, levels=myvar))
iters        <- combYI %>% 
  filter(stk == mystk, qname %in% c(myvar),  .id %in% scenarios, year %in% 2000:2021, iter <= 5) %>% 
  mutate(.id = factor(.id, levels=scenarios)) %>% 
  group_by(stk, .id, iter) %>% 
  mutate(perc_diff = (data-lag(data))/lag(data))

df %>% 
  ggplot(aes(x=year, y=median)) +
  theme_bw() +
  theme(panel.grid.minor = element_blank(),  
        panel.border     = element_rect(linetype = 1, colour="black", fill=NA))+

  # vline at 2010
  geom_vline(xintercept = 2010, colour="darkgray", linewidth=1) +

  # iters
  geom_line(data=iters,
            aes(y=data, group=iter),
            linewidth=0.4, colour="darkgray") +

  # median and ribbon
  geom_line(aes(y=median), 
            linewidth=1, colour="red") +
  geom_ribbon(aes(xmin=year, xmax=year, ymin=lowq, ymax=uppq), 
              alpha=0.2, fill="red") +

  # ICES
  geom_line(data= ices, aes(year,median), 
            col="blue", linewidth=1, linetype=2, inherit.aes=FALSE) +

  # refpoints
  geom_line(data=rp, 
            aes(x=year, y=data, group=reftype, col=reftype, linetype=reftype),
            linewidth=1) +
  scale_linetype_manual(values = c("longdash","dashed"), breaks = c("lim", "msy") ) +
  scale_colour_manual  (values = c("black","darkgray"), breaks = c("lim", "msy") ) +

  expand_limits(y=0) +
  facet_grid(qname ~ .id, scales="free_y")

@@

**Figure `r mystk`.3**  `r mystkname`. Comparison of different HCR’s with and without TAC bounds

@@

#### `r mystkname`: comparing HCR’s with and without a cap on TAC bounds

HCR1 bnd
HCR1 bnd cap
HCR2 bnd
HCR2 bnd cap

```{r, runWhb3, echo=FALSE, fig.asp=1.15, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}

myvar     <- c("catch","cCum", "ssb","fbar")
scenarios <- c("HCR1 bnd", "HCR1 bnd cap", "HCR2 bnd", "HCR2 bnd cap")
rp        <- refs %>% 
  filter(stk == mystk, refpoint %in% c("fmsy", "blim", "msybtrigger")) %>% 
    mutate(qname = factor(qname, levels=myvar))
ices      <- combY %>% 
  filter(stk == mystk, qname %in% c(myvar),  .id %in% "ICES",  year %in% 2000:2021) %>% 
  dplyr::select(- .id) %>% 
  mutate(qname = factor(qname, levels=myvar))
df        <- combY %>% 
  filter(stk == mystk, qname %in% c(myvar),  .id %in% scenarios, year %in% 2000:2021) %>% 
  mutate(.id = factor(.id, levels=scenarios)) %>% 
  mutate(qname = factor(qname, levels=myvar))
iters        <- combYI %>% 
  filter(stk == mystk, qname %in% c(myvar),  .id %in% scenarios, year %in% 2000:2021, iter <= 5) %>% 
  mutate(.id = factor(.id, levels=scenarios)) %>% 
  group_by(stk, .id, iter) %>% 
  mutate(perc_diff = (data-lag(data))/lag(data))

df %>% 
  ggplot(aes(x=year, y=median)) +
  theme_bw() +
  theme(panel.grid.minor = element_blank(),  
        panel.border     = element_rect(linetype = 1, colour="black", fill=NA))+

  # vline at 2010
  geom_vline(xintercept = 2010, colour="darkgray", linewidth=1) +

  # iters
  geom_line(data=iters,
            aes(y=data, group=iter),
            linewidth=0.4, colour="darkgray") +

  # median and ribbon
  geom_line(aes(y=median), 
            linewidth=1, colour="red") +
  geom_ribbon(aes(xmin=year, xmax=year, ymin=lowq, ymax=uppq), 
              alpha=0.2, fill="red") +
  # ICES
  geom_line(data= ices, aes(year,median), 
            col="blue", linewidth=1, linetype=2, inherit.aes=FALSE) +

  # refpoints
  geom_line(data=rp, 
            aes(x=year, y=data, group=reftype, col=reftype, linetype=reftype),
            linewidth=1) +
  scale_linetype_manual(values = c("longdash","dashed"), breaks = c("lim", "msy") ) +
  scale_colour_manual  (values = c("black","darkgray"), breaks = c("lim", "msy") ) +

  expand_limits(y=0) +
  facet_grid(qname ~ .id, scales="free_y")

@@

**Figure `r mystk`.4**  `r mystkname`. Comparison of different HCR’s with and without a cap on TAC bounds

@@

#### `r mystkname`: comparing HCR’s with and without implementation error (without bounds on TAC change)

HCR0
HCR1 10%
HCR1 20%
HCR1 30%
HCR1 10% recent
HCR1 10% random recent

```{r, runWhb4, echo=FALSE, fig.asp=1.15, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}

myvar     <- c("catch","cCum", "ssb","fbar")
scenarios <- c("HCR1", "HCR1 +10%", "HCR1 +20%", "HCR1 +30%", "HCR1 +10% recent", 
               "HCR1 +random recent")
rp        <- refs %>% 
  filter(stk == mystk, refpoint %in% c("fmsy", "blim", "msybtrigger")) %>% 
  mutate(qname = factor(qname, levels=myvar))
ices      <- combY %>% 
  filter(stk == mystk, qname %in% c(myvar),  .id %in% "ICES",  year %in% 2000:2021) %>% 
  dplyr::select(- .id) %>% 
  mutate(qname = factor(qname, levels=myvar))
df        <- combY %>% 
  filter(stk == mystk, qname %in% c(myvar),  .id %in% scenarios, year %in% 2000:2021) %>% 
  mutate(.id = factor(.id, levels=scenarios)) %>% 
  mutate(qname = factor(qname, levels=myvar))
iters        <- combYI %>% 
  filter(stk == mystk, qname %in% c(myvar),  .id %in% scenarios, year %in% 2000:2021, iter <= 5) %>% 
  mutate(.id = factor(.id, levels=scenarios)) %>% 
  group_by(stk, .id, iter) %>% 
  mutate(perc_diff = (data-lag(data))/lag(data))

df %>% 
  ggplot(aes(x=year, y=median)) +
  theme_bw() +
  theme(panel.grid.minor = element_blank(),  
        panel.border     = element_rect(linetype = 1, colour="black", fill=NA))+

  # vline at 2010
  geom_vline(xintercept = 2010, colour="darkgray", linewidth=1) +

  # iters
  geom_line(data=iters,
            aes(y=data, group=iter),
            linewidth=0.4, colour="darkgray") +

  # median and ribbon
  geom_line(aes(y=median), 
            linewidth=1, colour="red") +
  geom_ribbon(aes(xmin=year, xmax=year, ymin=lowq, ymax=uppq), 
              alpha=0.2, fill="red") +

  # ICES
  geom_line(data= ices, aes(year,median), 
            col="blue", linewidth=1, linetype=2, inherit.aes=FALSE) +

  # refpoints
  geom_line(data=rp, 
            aes(x=year, y=data, group=reftype, col=reftype, linetype=reftype),
            linewidth=1) +
  scale_linetype_manual(values = c("longdash","dashed"), breaks = c("lim", "msy") ) +
  scale_colour_manual  (values = c("black","darkgray"), breaks = c("lim", "msy") ) +

  expand_limits(y=0) +
  facet_grid(qname ~ .id, scales="free_y")

@@

**Figure `r mystk`.5**  `r mystkname`. Comparison of different HCR’s with and without implementation error (without bounds on TAC change)

@@

#### `r mystkname`: comparing HCR’s with and without implementation error in combination with bounds on TAC change

HCR1 bnd
HCR1 10% bnd
HCR1 10% recent bnd
HCR1 10% random bnd
HCR1 10% random recent bnd

```{r, runWhb5, echo=FALSE, fig.asp=1.15, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}

myvar     <- c("catch","cCum", "ssb","fbar")
scenarios <- c("HCR1 bnd", 
               "HCR1 +10% bnd",
               "HCR1 +10% bnd recent",
               "HCR1 +random bnd", 
               "HCR1 +random bnd recent")
rp        <- refs %>% 
  filter(stk == mystk, refpoint %in% c("fmsy", "blim", "msybtrigger")) %>% 
  mutate(qname = factor(qname, levels=myvar))
ices      <- combY %>% 
  filter(stk == mystk, qname %in% c(myvar),  .id %in% "ICES",  year %in% 2000:2021) %>% 
  dplyr::select(- .id) %>% 
  mutate(qname = factor(qname, levels=myvar))
df        <- combY %>% 
  filter(stk == mystk, qname %in% c(myvar),  .id %in% scenarios, year %in% 2000:2021) %>% 
  mutate(.id = factor(.id, levels=scenarios)) %>% 
  mutate(qname = factor(qname, levels=myvar))
iters        <- combYI %>% 
  filter(stk == mystk, qname %in% c(myvar),  .id %in% scenarios, year %in% 2000:2021, iter <= 5) %>% 
  mutate(.id = factor(.id, levels=scenarios)) %>% 
  group_by(stk, .id, iter) %>% 
  mutate(perc_diff = (data-lag(data))/lag(data))

df %>% 
  ggplot(aes(x=year, y=median)) +
  theme_bw() +
  theme(panel.grid.minor = element_blank(),  
        panel.border     = element_rect(linetype = 1, colour="black", fill=NA))+

  # vline at 2010
  geom_vline(xintercept = 2010, colour="darkgray", linewidth=1) +

  # iters
  geom_line(data=iters,
            aes(y=data, group=iter),
            linewidth=0.4, colour="darkgray") +

  # median and ribbon
  geom_line(aes(y=median), 
            linewidth=1, colour="red") +
  geom_ribbon(aes(xmin=year, xmax=year, ymin=lowq, ymax=uppq), 
              alpha=0.2, fill="red") +

  # ICES
  geom_line(data= ices, aes(year,median), 
            col="blue", linewidth=1, linetype=2, inherit.aes=FALSE) +

  # refpoints
  geom_line(data=rp, 
            aes(x=year, y=data, group=reftype, col=reftype, linetype=reftype),
            linewidth=1) +
  scale_linetype_manual(values = c("longdash","dashed"), breaks = c("lim", "msy") ) +
  scale_colour_manual  (values = c("black","darkgray"), breaks = c("lim", "msy") ) +

  expand_limits(y=0) +
  facet_grid(qname ~ .id, scales="free_y")

@@

**Figure `r mystk`.6**  `r mystkname`. Comparison of different HCR’s with and without implementation error in combination with bounds on TAC change

@@

#### `r mystkname`: checking the working of different constraints

```{r, whbcheck, echo=FALSE, fig.asp=.5, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}

myvar     <- c("ssb")
# scenarios <- c("ICES", "HCR1 bnd", "HCR1 +10%", "HCR1 +10% bnd", "HCR1 +10% bnd cap")
scenarios <- c("HCR1", "HCR1 bnd", "HCR2", "HCR2 bnd")
#scenarios <- c("ICES", "HCR1 bnd", "HCR1 +10% bnd", "HCR1 +20% bnd", "HCR1 +30% bnd")
rp        <- refs %>% filter(stk == mystk, refpoint %in% c("msybtrigger", "blim"))
ices      <- combY %>% filter(stk == mystk, qname %in% c(myvar),  .id %in% "ICES",  year %in% 2000:2021) %>% dplyr::select(- .id)
df        <- combY %>% filter(stk == mystk, qname %in% c(myvar),  .id %in% scenarios, year %in% 2000:2021) %>% 
  mutate(.id = factor(.id, levels=scenarios)) %>% 
  left_join(dplyr::select(rp,
                          stk, year, qname, refpoint, reftype, rp=data), 
            by=c("stk", "year", "qname")) %>% 
  mutate(trigger = ifelse(median >= rp, "above","below"))
iters        <- combYI %>% 
  filter(stk == mystk, qname %in% c(myvar),  .id %in% scenarios, year %in% 2000:2021, iter <= 10) %>% 
  mutate(.id = factor(.id, levels=scenarios)) %>% 
  left_join(dplyr::select(rp,
                          stk, year, qname, refpoint, reftype, rp=data), 
            by=c("stk", "year", "qname")) %>% 
  mutate(trigger = ifelse(data >= rp, "above","below"))
df %>% 
  ggplot(aes(x=year, y=median)) +
  theme_bw() +
  theme(panel.grid.minor = element_blank(),  
        panel.border     = element_rect(linetype = 1, colour="black", fill=NA))+
  geom_line(data=iters,
            aes(y=data, group=iter),
            linewidth=0.4, colour="darkgray") +
  geom_point(data=filter(iters,
                         trigger=="above"),
            aes(y=data, group=iter),
            size=1.0, colour="black") +

  geom_line(linewidth=1, colour="red") +
  geom_point(data=filter(df,
                        trigger=="above"),
            colour="red", size=2) +
    # refpoints
  geom_line(data=rp, 
            aes(x=year, y=data, group=reftype, col=reftype, linetype=reftype),
            linewidth=1) +
  scale_linetype_manual(values = c("longdash","dashed"), breaks = c("lim", "msy") ) +
  scale_colour_manual  (values = c("black","darkgray"), breaks = c("lim", "msy") ) +
  labs(y="ssb (tonnes)") +
  expand_limits(y=0) +
  facet_grid(qname ~ .id, scales="free_y")

myvar     <- c("catch")
# scenarios <- c("ICES", "HCR1 bnd", "HCR1 +10%", "HCR1 +10% bnd", "HCR1 +10% bnd cap")
# scenarios <- c("ICES", "HCR1 bnd", "HCR1 +10% bnd", "HCR1 +20% bnd", "HCR1 +30% bnd")
rp        <- refs %>% 
  filter(stk == mystk, refpoint %in% c("fmsy", "blim", "msybtrigger")) %>% 
  mutate(qname = factor(qname, levels=myvar))
ices      <- combY %>% 
  filter(stk == mystk, qname %in% c(myvar),  .id %in% "ICES",  year %in% 2000:2021) %>% 
  dplyr::select(- .id) %>% 
  mutate(qname = factor(qname, levels=myvar))
df        <- combY %>% 
  filter(stk == mystk, qname %in% c(myvar),  .id %in% scenarios, year %in% 2000:2021) %>% 
  mutate(.id = factor(.id, levels=scenarios)) %>% 
  mutate(qname = factor(qname, levels=myvar)) %>% 
  group_by(stk, .id) %>% 
  mutate(perc_diff = (median-lag(median))/lag(median))

iters        <- combYI %>% 
  filter(stk == mystk, qname %in% c(myvar),  .id %in% scenarios, year %in% 2000:2021, iter <= 5) %>% 
  mutate(.id = factor(.id, levels=scenarios)) %>% 
  group_by(stk, .id, iter) %>% 
  mutate(perc_diff = (data-lag(data))/lag(data))

df %>% 
  ggplot(aes(x=year, y=perc_diff)) +
  theme_bw() +
  theme(panel.grid.minor = element_blank(),  
        panel.border     = element_rect(linetype = 1, colour="black", fill=NA))+

  geom_line(data=iters,
            aes(x=year, y=perc_diff, group=iter),
            linewidth=.4, colour="darkgray", inherit.aes = FALSE) +

  geom_line(linewidth=1, colour="red") +

  geom_hline(yintercept = 0.25, colour="black", linewidth=1, linetype=2) +
  geom_hline(yintercept = -0.2, colour="black", linewidth=1, linetype=2) +
  expand_limits(y=0) +
  scale_y_continuous(labels=scales::percent, limits=c(-2,2)) +
  facet_grid(qname ~ .id, scales="free_y")

@@

**Figure `r mystk`.6**  `r mystkname`. Comparison of the working of constraints on TAC change relative to stock size and reference point MSY Btrigger. Black dots indicate estimated stock size on individual runs above MSY Btrigger (i.e. TAC contstraint applies). Red dots indicate median stock size above MSY Btrigger.

@@

#### `r mystkname`: comparing all scenario's on median catch, IAV and AAV

```{r, whbsmry, echo=FALSE, fig.asp=1.2, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}

myvar     <- c("catch.50", "iav.50", "aav")
scenarios <- c("ICES","Fmsy","HCR1","HCR1 bnd","HCR1 bnd cap", "HCR1 +10% bnd","HCR1 +10% bnd cap",
               "HCR1 +10% bnd recent","HCR1 +random bnd",",HCR1 +random bnd recent","HCR1 +10%",
               "HCR1 +10% recent","HCR1 +random","HCR1 +random recent","HCR1 +20% bnd", 
               "HCR1 +20% bnd cap","HCR1 +20% bnd recent","HCR1 +20%","HCR1 +20% recent",
               "HCR1 +30% bnd","HCR1 +30% bnd cap","HCR1 +30% bnd recent","HCR1 +30%",
               "HCR1 +30% recent",
               "HCR2","HCR2 bnd","HCR2 bnd cap","HCR2 +10% bnd","HCR2 +10% bnd cap",
               "HCR2 +10% bnd recent","HCR2 +random bnd","HCR2 +random bnd recent","HCR2 +10%",
               "HCR2 +10% recent","HCR2 +random","HCR2 +random recent","HCR2 +20% bnd",
               "HCR2 +20% bnd cap","HCR2 +20% bnd recent","HCR2 +20%","HCR2 +20% recent",
               "HCR2 +30% bnd","HCR2 +30% bnd cap","HCR2 +30% bnd recent","HCR2 +30%",
               "HCR2 +30% recent")

df        <- combI %>% 
  tidyr::pivot_longer(names_to = "variable", values_to = "data", myvar) %>% 
  filter(stk == mystk,  .id %in% scenarios) %>% 
  mutate(.id = factor(.id, levels=scenarios)) %>% 
  mutate(variable = factor(variable, levels=myvar))

calc_boxplot_stat <- function(x) {
  coef <- 1.5
  n <- sum(!is.na(x))
  # calculate quantiles
  stats <- quantile(x, probs = c(0.0, 0.25, 0.5, 0.75, 1.0))
  names(stats) <- c("ymin", "lower", "middle", "upper", "ymax")
  iqr <- diff(stats[c(2, 4)])
  # set whiskers
  outliers <- x < (stats[2] - coef * iqr) | x > (stats[4] + coef * iqr)
  if (any(outliers)) {
    stats[c(1, 5)] <- range(c(stats[2:4], x[!outliers]), na.rm = TRUE)
  }
  return(stats)
}

df %>% 
  ggplot(aes(x=.id, y=data)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  stat_summary(fun.data = calc_boxplot_stat, geom="boxplot") +
  expand_limits(y=0) +
  # scale_y_continuous(limits = quantile(t$data, c(0.1, 0.9))) +
  labs(x="", y="") +
  facet_grid(variable~., scales="free", space="free_x")

@@

**Figure `r mystk`.7**  `r mystkname`. Comparison of all scenario's on median catch, median IAV (Interannual variability in catch) and AAV (Average Annual Variability in catch)

```{r smry2, eval=FALSE, fig.align="center", fig.asp=1.2, message=FALSE, warning=FALSE, cache=FALSE, include=FALSE, whbsmry, echo=FALSE}

>>>>

library(captioner)     # captions
tab_nums <- captioner::captioner(prefix = "Table", levels=1, type=c("n"), infix=".")
fig_nums <- captioner::captioner(prefix = "Figure", levels=1, type=c("n"), infix=".")

source("../R/dd.R")
source("../R/pm.R")

source("../R/get_dropbox.r")
dropboxdir<-try(file.path(get_dropbox(), "pelagics"))

@@

# mystk     <- "mac"; mystkname <- "Northeast Atlantic mackerel"; mylatin <- "Scombrus scomber"
mystk     <- "whb"; mystkname <- "Blue whiting";                mylatin <- "Micromesistius poutassou" 
maxyear <- 2050

load(file=file.path(dropboxdir,paste0("data/om/",mystk,"matPar.RData")))

# ------------------------------------------------------------------------------
# Operating model
# ------------------------------------------------------------------------------
figuresdir <- file.path(dropboxdir, "results", mystk, "figures")
tablesdir  <- file.path(dropboxdir, "results", mystk, "tables")

@@

<!-- =============================================================================================================== -->
<!-- ## Mackerel report                                                                                              -->
<!-- =============================================================================================================== -->

**Evaluating impacts of Density Dependent processes on `r mystkname` via MSE simulations**

**L. Kell, M.A. Pastoors**

@@

Commissioned by: University of Copenhagen (Henrik Sparholt)

**Executive summary**

[summary here]

<!-- =============================================================================================================== -->
<!-- Section 1: introduction                                                                                         -->
<!-- =============================================================================================================== -->

```{r}

# captioner::bump(fig_nums, level=1)
# fig_nums(name    = "section1_fig", caption = "section1_fig", level = 1, display = FALSE)
# tab_nums(name    = "section1_tab", caption = "section1_tab", level = 1, display = FALSE)

@@

# Introduction

The purpose of this work is to support the project “Fmsy-project for six high profile fish stocks”. The project focuses on the current management approach for fish stocks based on single-species Fmsy Values where the only density-dependent process that is included is recruitment. The project includes three other density-dependent population dynamics: growth (weight at age), sexual maturity and natural mortality.

This report deals with the evaluation of density dependent processes for `r mystkname` using Management Strategy Evaluation (MSE). Steps in the analysis:

1. Conditioning and summaries based on the 2022 ICES assessment (section 2)
2. Estimating Density Dependent effects based on equilibrium analyses and explore DD before setting up the OMs (section 3)
3. Setting up the Operating Models (OM) (section 4)
4. Running MSE's (section 5)

Four alternatives OMs have been included in the analyses:

1. No density dependent processes
2. DD in growth
3. DD in growth and maturity
4. DD in growth, maturity and natural mortality

All MSE's have been developed within the FLR framework (https://flr-project.org/) and all the code is stored on github (https://github.com/martinpastoors/ddmse).

All data and results are available on DropBox (https://www.dropbox.com/sh/k87x074xqw7sgae/AAAmF980iR2n24wTkqhOhnpfa?dl=0)

<!-- =============================================================================================================== -->
<!-- Section 2: conditioning and summaries of ICES assessment                                                        -->
<!-- =============================================================================================================== -->

# Conditioning and summaries based on the 2022 ICES assessment

```{r}

# fig_nums(name    = "section2_fig", caption = "section2_fig", level = 1, display = FALSE)
# tab_nums(name    = "section2_tab", caption = "section2_tab", level = 1, display = FALSE)

**Weight at age**

[text here]

```{r, echo=FALSE, fig.width=10, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}

fig_nums(
  name    = "cond_waa", level = 1, display = FALSE,
  caption = paste0(mystkname, ". Weight at age by year and cohort"))

knitr::include_graphics(path=file.path(figuresdir, paste(mystk, "waa.jpg", sep="_")))

_`r fig_nums("cond_waa")`_

@@

**Maturity at age**

[text here]

```{r, echo=FALSE, fig.width=10, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}

fig_nums(
  name    = "cond_mat", level = 1, display = FALSE,
  caption = paste0(mystkname, ". Maturity at age by year and cohort"))

knitr::include_graphics(path=file.path(figuresdir, paste(mystk, "mat.jpg", sep="_")))

@@

_`r fig_nums("cond_mat")`_

**Natural mortality at age (fixed)**

[text here]

```{r, echo=FALSE, fig.width=10, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}

fig_nums(
  name    = "cond_natmor", level = 1, display = FALSE,
  caption = paste0(mystkname, ". Natural mortality at age by year and cohort (all ages and years the same M)"))

knitr::include_graphics(path=file.path(figuresdir, paste(mystk, "m.jpg", sep="_")))

@@

_`r fig_nums("cond_natmor")`_

@@

**Density dependence in Weight at age vs. Total biomass**

[text here]

```{r, echo=FALSE, fig.width=10, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}

fig_nums(
  name    = "cond_ddwaa", level = 1, display = FALSE,
  caption = paste0(mystkname, ". Density dependence and weight at age."))

knitr::include_graphics(path=file.path(figuresdir, paste(mystk, "dd_waa.jpg", sep="_")))

@@

_`r fig_nums("cond_ddwaa")`_

@@

**GAMM at age**

[text here]

```{r, echo=FALSE, fig.width=10, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}

fig_nums(
  name    = "cond_gammwaa", level = 1, display = FALSE,
  caption = paste0(mystkname, ". GAMM analysis of density dependence and weight at age."))

knitr::include_graphics(path=file.path(figuresdir, paste(mystk, "gamm.jpg", sep="_")))

@@

_`r fig_nums("cond_gammwaa")`_

@@

**GAMM summary**

[text here]

```{r, echo=FALSE, fig.width=10, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}

tab_nums(
  name    = "cond_gammsummary", level = 1, display = FALSE,
  caption = paste0(mystkname, ". GAMM summary of density dependence and weight at age."))

cat(readLines(file.path(tablesdir, paste(mystk, "gamm summary.txt", sep="_"))), sep = '\n')

@@

_`r tab_nums("cond_gammsummary")`_

@@

**GAMM summary**

[text here]

```{r, echo=FALSE, fig.width=10, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}

tab_nums(
  name    = "cond_gammcoefficients", level = 1, display = FALSE,
  caption = paste0(mystkname, ". GAMM coefficients of density dependence and weight at age."))

cat(readLines(file.path(tablesdir, paste(mystk, "gamm coefficients.txt", sep="_"))), sep = '\n')

@@

_`r tab_nums("cond_gammcoefficients")`_

@@

**LM of density dependence in weight at age**

[text here]

```{r, echo=FALSE, fig.width=10, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}

fig_nums(
  name    = "cond_lmwaa", level = 1, display = FALSE,
  caption = paste0(mystkname, ". LM analysis of density dependence and weight at age."))

knitr::include_graphics(path=file.path(figuresdir, paste(mystk, "lm_waa.jpg", sep="_")))

@@

_`r fig_nums("cond_lmwaa")`_

@@

**LM summary**

[text here]

```{r, echo=FALSE, fig.width=10, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}

tab_nums(
  name    = "cond_lmsummary", level = 1, display = FALSE,
  caption = paste0(mystkname, ". LM summary of density dependence and weight at age."))

cat(readLines(file.path(tablesdir, paste(mystk, "lm summary.txt", sep="_"))), sep = '\n')

@@

_`r tab_nums("cond_gammcoefficients")`_

**DD analysis of maturity at age**

[text here]

```{r, echo=FALSE, fig.width=10, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}

fig_nums(
  name    = "cond_ddmat", level = 1, display = FALSE,
  caption = paste0(mystkname, ". Analysis of density dependence and maturity at age."))

knitr::include_graphics(path=file.path(figuresdir, paste(mystk, "dd_mat.jpg", sep="_")))
